---
title: "a05"
author: "Chirar, Layla"
editor: visual
---

## Introduction

This data is from the EDI Data Repository: Sacramento-San Joaquin Delta Socioecological Monitoring

## Analysis

```{r}
#Loading packages
library(readr)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(scales) # scale functions for visualization
```

```{r} unnamed chunk?}
#needed to install the following packages
install.packages("janitor")
install.packages("DT")
install.packages("leaflet")
install.packages("viridis")
```

```{r}
#loading newly installed packages
library(janitor)
library(leaflet)
library(DT)
library(viridis)
```

```{r}
#loads data
delta_visits_raw <- read_csv(here::here("data/raw/Socioecological_monitoring_data.csv"))

```

```{r}
#gives names of columns
colnames(delta_visits_raw)
```

```{r}
#shows how the data is organized
glimpse(delta_visits_raw)
```

```{r}
#gives the date range for the dataset
range(delta_visits_raw$Date)
```

```{r}
#gives the unique values in Time_of_Day
unique(delta_visits_raw$Time_of_Day)
```

```{r}
#creates new variable where all names have the same format
delta_visits <- delta_visits_raw %>% 
    janitor::clean_names()
```

```{r}
#makes data tidier with long format
visits_long <- delta_visits %>%
  pivot_longer(cols = c(sm_boat, med_boat, lrg_boat, bank_angler, scientist, cars),
                 names_to = "visitor_type",
                 values_to = "quantity") %>%
    rename(restore_loc = eco_restore_approximate_location) %>% 
    select(-notes)

## Checks outcome
head(visits_long)
```

```{r}
daily_visits_loc <- visits_long %>% 
    group_by(restore_loc, date, visitor_type) %>% # splits dataset into 3 groups
    summarise(daily_visits = sum(quantity)) #adds all quanities and creates new variable
    

```

```{r}
#re orders the categories
daily_visits_totals <- daily_visits_loc %>% 
    group_by(restore_loc) %>% #all of the sam elcoation are grouped together
    mutate(n = sum(daily_visits)) %>% #adds daily visits for each location
    ungroup() %>%
    mutate(restore_loc = fct_reorder(restore_loc, n)) #reorders based on total visits

head(daily_visits_totals)
```

```{r}
ggplot(data = daily_visits_totals, #creates plot using the daily_visit_totals
       aes(x = daily_visits, y = restore_loc, #selects the variable for each axis
           fill = visitor_type)) +
    geom_col() +
    labs(x = "Number of Visits", #labels everything
         y = "Restoration Location", #labels the axis
         fill = "Type of Visitor", #fills the bar based on the variable
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period") +
    scale_x_continuous(breaks = seq(0,120, 20), expand = c(0,0)) +
    theme_bw()
```

```{r}
locations <- visits_long %>% 
    distinct(restore_loc, .keep_all = TRUE) %>% #uses rows from restore_loc
    select(restore_loc, latitude, longitude) #keeps columns restore_loc, latitude, longitude
```

```{r}
leaflet(locations) %>%
    addTiles() %>% #adds tiles to map
    addMarkers( #adds markers for locations with pop up with location from restore_loc
        lng = ~ longitude,
        lat = ~ latitude,
        popup = ~ restore_loc
    )
```

## Discussion

The bar plot shows each location's total number of visitors in descending order and also splits the bars up to show how many of the total visitors for each location belong to different groups of types of visitors. Med boats appeared to be the most abundant visitor type across all locations.
